% ============= CHAPTER 1: INTRODUCTION =============
\chapter{Introduction}
\label{ch:chapter1}

\section{Introduction}

This project focuses on the design, development, and implementation of the Threat Intelligence-Based Security Assessment (TIBSA) platform, a unified full-stack cybersecurity application. The platform is engineered to seamlessly integrate threat modeling, automated vulnerability scanning, threat intelligence ingestion, malware sandboxing analysis, and incident correlation. By combining the predictive analysis of threat modeling with the empirical validation of automated penetration testing and real-time threat intelligence feeds, TIBSA delivers a context-aware and risk-driven security posture assessment.

Unlike traditional security workflows where threat analysis, scanning, and incident investigation operate in isolated silos, TIBSA provides a unified workspace. In this workspace, identified architectural threats directly generate targeted security scanning jobs, while scan findings dynamically feed back to update the system's active threat model database, adjusting severity levels and recalculating the overall risk score. Additionally, the platform integrates advanced malware and URL analyzers (leveraging static and dynamic sandboxing, external intelligence APIs, and local machine learning models), a multi-stage incident investigation orchestrator (generating interactive attack chains and log timelines), and an interactive AI security chatbot. This unified framework bridges the gap between design-time security analysis and run-time security testing, allowing organizations to prioritize business-critical risks, simulate real-world adversary tactics, and dynamically adapt to an ever-evolving threat landscape.

\section{Problem Statement}

In many organizations, security analysis, vulnerability testing, malware research, and threat intelligence operate as fragmented, disconnected processes. This lack of integration leads to several critical security management issues:

\begin{itemize}
    \item \textbf{Isolated Threat Modeling:} Traditional threat modeling (such as STRIDE) identifies potential architectural weaknesses during the design phase but lacks direct empirical validation. There is no automated link to verify if these theoretical threats translate into actual, exploitable vulnerabilities in the running system.
    \item \textbf{Context-Unaware Security Scanning:} Automated penetration testing and vulnerability scanners discover weaknesses but do not leverage threat intelligence or design-phase threat models. Consequently, security teams waste resources remediating low-impact findings instead of focusing on critical business risks and active attack vectors.
    \item \textbf{Siloed Malware and URL Analysis:} Suspicious files and URLs are typically analyzed using standalone third-party websites or disconnected local tools. The results and indicators of compromise (IOCs) are not automatically correlated with the system's threat model or historical scan data.
    \item \textbf{Delayed Incident Investigation and Correlation:} When security events occur, analysts must manually reconstruct attack vectors by linking logs, scanning findings, and intelligence feeds. This lack of centralized event correlation slows down incident response and increases system exposure times.
\end{itemize}

This disconnection leads to delayed threat detection, misprioritized vulnerabilities, and repeated exposures---leaving systems vulnerable to preventable attacks. Therefore, a unified, interactive platform is required to integrate Threat Modeling, Penetration Testing, Malware Analysis, and Incident Investigation, creating a continuous, intelligence-driven, and adaptive security process.

\section{Motivation}

As cyberattacks become more automated, multi-staged, and complex, traditional point-in-time assessments (like annual penetration tests or static threat diagrams) are insufficient to ensure ongoing protection. The integration of threat modeling, automated scanning, malware analysis, and real-time threat intelligence into a unified platform provides several key advantages:

\begin{itemize}
    \item \textbf{Risk-Centric Prioritization:} Empowers security teams to prioritize vulnerabilities based on realistic business risk scoring (inherent vs. residual risk) and compliance guidelines (OWASP ASVS), rather than relying on generic severity ratings.
    \item \textbf{Continuous Security Validation:} Establishes a direct loop where threats identified in the system model are translated into target test cases, and actual scan results dynamically update the state and confidence level of modeled threats.
    \item \textbf{Proactive Threat Intelligence Integration:} Incorporates real-world adversary behavioral models (CAPEC attack patterns mapped to CWEs) and file reputation checks (VirusTotal API) to identify and analyze active threats.
    \item \textbf{Rapid Incident Response:} Automates the correlation of security incidents and host activities, mapping them into interactive node-based attack chains to speed up diagnostic investigations.
\end{itemize}

For cybersecurity professionals and system developers, this platform provides a unified workspace that bridges the gap between predictive design-level threat modeling and active operational validation, aligning security processes with modern agile DevSecOps workflows.

\section{Project Objectives}

The primary objectives of this project are:

\begin{enumerate}
    \item \textbf{Develop a Unified Cybersecurity Platform:} Implement a robust full-stack web application consisting of a Next.js (TypeScript) frontend, a FastAPI (Python) backend, and a Supabase database and authentication infrastructure.
    \item \textbf{Build an Advanced Threat Modeling Engine:} Develop an automated STRIDE-based threat modeling service enriched with CAPEC attack patterns, mapped to OWASP ASVS compliance controls, and integrated with dynamic risk-scoring and heatmap generators.
    \item \textbf{Implement a Vulnerability Scanning Engine:} Create an automated penetration testing scanner to check targets for configuration weaknesses, security headers, cookie properties, and directory paths, feeding results back to the system.
    \item \textbf{Establish a Continuous Posture Feedback Loop:} Design a database synchronization model where the vulnerability scanner's findings automatically update the status (e.g., potential vs. verified) of threats in the model.
    \item \textbf{Incorporate a Malware and URL Analysis Layer:} Implement an analysis module capable of scanning suspicious files and URLs, integrating external API reputations, and running local machine learning classifiers to predict phishing probability.
    \item \textbf{Orchestrate Security Incident Investigations:} Create a correlation engine that aggregates scan logs and security event feeds to map potential multi-stage threat movements and visualize them as attack chains.
    \item \textbf{Integrate an AI Security Chatbot:} Embed an LLM-powered interactive security assistant to guide analysts through threat modeling inputs, scan diagnostics, and mitigation rules.
\end{enumerate}

\section{Expected Solutions}

The TIBSA platform delivers a complete set of integrated solutions to address modern security assessment and incident correlation requirements:

\subsection{Interactive Full-Stack Web Platform}
A web application built using Next.js for a responsive, modern frontend and FastAPI for a lightweight, high-performance backend, utilizing Supabase for secure multi-tenant user authentication, database management, and role-based access control (Admin/User).

\subsection{Automated STRIDE Threat Modeling Engine}
A modeling module that processes system metadata, trust boundaries, and asset inputs to generate threat logs mapped to the CAPEC database (including CWE references) and OWASP ASVS compliance checks. It calculates inherent and residual risk scores using weighted formulas and visualizes them on an interactive risk matrix heatmap.

\subsection{Automated Vulnerability Scanner (Penetration Testing)}
A scanning component that conducts automated penetration testing on target domains. The scanner fingerprints web technologies, brute-forces common directory paths, evaluates security-related HTTP headers, and analyzes cookie configurations (such as HttpOnly and Secure flags).

\subsection{Continuous Feedback and Posture Update Loop}
A database synchronization workflow where penetration testing findings are automatically linked back to matching categories in the threat model, updating threat statuses from potential to verified, adjusting the residual risk score, and reflecting updates in the system dashboard.

\subsection{Malware and URL Threat Analysis Layer}
An analysis dashboard where users upload suspicious files or input URLs/hashes. The platform performs reputational audits via VirusTotal API integration, extracts indicators of compromise (IOCs), and executes local machine learning models (trained on URL phishing indicators) to classify malice probabilities.

\subsection{Incident and SOC Event Correlation Dashboard}
A monitoring center that processes system events, scanner findings, and network records. It automatically correlates these occurrences to reconstruct multi-stage adversary movements, presenting them as interactive node graphs (attack chains) and historical chronological event feeds.

\subsection{Interactive Security AI Chatbot}
An embedded virtual security consultant powered by a large language model. The chatbot helps security analysts query threat rules, write customized security policies, interpret scan logs, and recommend specific mitigation steps.

\subsection{Comprehensive Reporting Engine}
A document generator that compiles threat models, scan findings, compliance gaps, and mitigations into formal, downloadable PDF and JSON security assessment reports.

\section{System Scope}

The scope of the TIBSA project covers the complete software development lifecycle, implementation, and deployment of a full-stack cybersecurity application. The components within scope are:

\begin{itemize}
    \item \textbf{Frontend Application:} A responsive user interface built using Next.js 16 (TypeScript) and styled with custom themes, CSS tokens, and interactive components. It includes dedicated dashboards for threat modeling, malware upload, vulnerability scanning, attack-chain visualization, and an AI chat interface.
    \item \textbf{Backend API Service:} A high-performance FastAPI (Python) application exposing REST endpoints for threat generation, scanning triggers, file reputation lookups, ML classification predictions, and report generation.
    \item \textbf{Database and Security Policy Layer:} A Supabase database storing project schemas (users, threat models, scan logs, vulnerabilities, findings, and events) and managing secure, role-based access permissions.
    \item \textbf{External Integrations:} Integration with the VirusTotal API for reputation services, as well as the CAPEC attack taxonomy and OWASP ASVS control database.
\end{itemize}

\section{Out of Scope}

To focus the project resources on core research and application logic, the following domains and capabilities are excluded from the implementation:

\begin{itemize}
    \item \textbf{Automated Source Code Patching:} The platform identifies vulnerabilities and provides step-by-step mitigation instructions, but it does not execute automated code modifications or directly patch source repositories.
    \item \textbf{Inline Network Traffic Filtering (IDS/IPS):} The application does not function as an active inline intrusion detection or prevention firewall (such as Snort or Suricata) operating at the network packet level.
    \item \textbf{Physical and Hardware-Level Penetration Testing:} Security testing focusing on hardware components, side-channel signal leaks, or physical entry audits (e.g., badge copying or locks) is excluded.
    \item \textbf{Threat Intelligence Feed Hosting:} The platform ingests external threat feeds and indicators of compromise (IOCs), but it does not host a public, global threat intelligence registry.
\end{itemize}

\section{Expected Outcomes}

The project is expected to yield the following deliverables:

\begin{itemize}
    \item A fully deployed, full-stack proof-of-concept platform (TIBSA) demonstrating the integration of threat modeling, scanning, and malware analysis.
    \item An automated workflow mapping system components to STRIDE categories, CAPEC identifiers, and OWASP ASVS guidelines.
    \item Visualized risk dashboards, dynamic heatmaps, and interactive attack chains that simplify security posturing and forensic audits.
    \item Downloadable security reports that detail identified vulnerabilities, active threat statuses, compliance scores, and actionable remediation steps.
\end{itemize}

\section{Conclusion}

The TIBSA platform establishes a unified, intelligence-driven framework that bridges the gap between design-stage threat prediction and operational-stage validation. By integrating STRIDE-based threat modeling, automated penetration testing scans, malware/URL reputation engines, and multi-stage event correlation into a responsive full-stack application, the platform empowers organizations to systematically identify vulnerabilities, assess overall risks, and protect their digital assets in an ever-evolving threat landscape.
