% ============= Abstract =============
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

Cybersecurity assessments in modern organizations often suffer from fragmented practices, where threat modeling, penetration testing, and continuous monitoring operate in isolation. This disconnection leads to inefficient risk management, redundant vulnerabilities, and delayed responses to evolving cyber threats.

This project proposes an intelligence-driven framework that integrates \textbf{Threat Modeling}, \textbf{Penetration Testing}, and \textbf{Threat Intelligence} to create a unified, risk-based security assessment process. By combining the predictive capabilities of threat modeling with the practical validation of penetration testing, the framework ensures that testing efforts are aligned with real business risks. The incorporation of threat intelligence enables adaptive and context-aware updates, allowing security assessments to evolve with the current threat landscape.

The proposed solution includes automated threat-to-test mapping, dynamic scope definition, centralized documentation, and continuous feedback loops for ongoing improvement. Expected outcomes include enhanced testing efficiency, better risk prioritization, and improved collaboration among security teams. Ultimately, this project aims to bridge the gap between design-time threat analysis and real-world defense validation, providing a modern, intelligence-driven approach to proactive cybersecurity management.
