% ============= REFERENCES =============
\renewcommand{\bibname}{List of References}
\begin{thebibliography}{99}
\addcontentsline{toc}{chapter}{List of References}

\bibitem{xu2012}
D. S. Xu, B. S. Chaparro, X. Ou, and P. Rao, ``Automated Security Test Generation with Formal Threat Models,'' \textit{IEEE Transactions on Dependable and Secure Computing (TDSC)}, 2012.

\bibitem{marback2013}
B. Marback, C. Doerr, P. Rao, B. S. Chaparro, and X. Ou, ``A Threat Model-Based Approach to Security Testing,'' \textit{Software: Practice and Experience}, 2013.

\bibitem{palanivel2014}
R. Palanivel and A. P. Selvadurai, ``Risk-Driven Security Testing Using Risk Analysis with Threat Modeling Approach,'' \textit{SpringerPlus}, 2014.

\bibitem{rahim2022}
F. A. Rahim, N. Jamil, Z. C. Cob, L. M. Sidek, and N. I. I. Sharizan, ``Risk Analysis of Water Grid Systems Using Threat Modeling,'' \textit{Journal of Physics: Conference Series}, vol. 2261, no. 1, p. 012015, 2022.

\bibitem{alozie2024}
C. E. Alozie, ``Threat Modeling in Health Care Sector,'' \textit{International Journal of Engineering and Modern Technology (IJEMT)}, vol. 10, no. 11, pp. 199--205, 2024.

\bibitem{granata2020}
D. Granata and M. Rak, ``Systematic analysis of automated threat modelling techniques: Comparison of open-source tools,'' \textit{Software Quality Journal}, vol. 32, pp. 125--161, 2024.

\bibitem{shin2022}
B. Shin, A. Elkins, L. Larson, M. Perez, and L. Cameron, ``Actionable Intelligence-Oriented Cyber Threat Modeling Framework,'' in \textit{Proc. 55th Hawaii Int. Conf. on System Sciences (HICSS)}, pp. 6782--6791, 2022.

\bibitem{dekker2023}
M. Dekker and L. Alevizos, ``A Threat-Intelligence Driven Methodology to Incorporate Uncertainty in Cyber Risk Analysis and Enhance Decision Making,'' \textit{Computers \& Security}, 2023.

\bibitem{smith2023}
J. Smith, R. Johnson, and L. Williams, ``Performing Ransomware Detection through Predictive Behavioral Mapping to Autonomous Threat Identification,'' \textit{IEEE Access}, vol. 10, pp. 14532--14548, 2022.

\bibitem{binsarhan2023}
B. Bin Sarhan and N. Altwaijry, ``Insider Threat Detection Using Machine Learning Approach,'' \textit{Applied Sciences}, vol. 13, no. 259, pp. 1--19, 2023.

\bibitem{alharbi2020}
M. Alharbi, E. Al-Shaer, and N. Almakhdhub, ``Actionable Intelligence-Oriented Cyber Threat Modeling Framework (TIME),'' in \textit{Proceedings of the 55th Hawaii International Conference on System Sciences (HICSS)}, pp. 6782--6791, 2022.

\bibitem{chu2021}
G. Chu, ``Automation of Penetration Testing,'' Ph.D. dissertation, Univ. of Liverpool, Liverpool, U.K., Sept. 2021.

\bibitem{huang2024}
J. Huang and Q. Zhu, ``PenHeal: A Two-Stage LLM Framework for Automated Pentesting and Optimal Remediation,'' \textit{arXiv preprint}, vol. abs/2407.17788, 2024.

\bibitem{chauhan2024}
K. Chauhan, ``Insider Threats Mitigation: Role of Penetration Testing,'' Department of Computer Science, University of Guelph, Guelph, Canada, 2024.

\bibitem{sanagana2023}
D. P. R. Sanagana, ``Mitigating Network Threats: Integrating Threat Modeling in Next-Generation Firewall Architecture,'' \textit{Journal of Science Technology and Research (JSTAR)}, vol. 4, no. 1, pp. 202--209, 2023.

\bibitem{maniraj2024}
S. P. Maniraj, C. S. Ranganathan, and S. Sekar, ``Securing web applications with OWASP ZAP for comprehensive security testing,'' \textit{Int. J. Adv. Sig. Img. Sci.}, vol. 10, no. 2, pp. 12--23, 2024.

% ============= Tools and Frameworks =============

\bibitem{nmap}
Nmap, ``Network Mapper,'' Wikipedia. [Online]. Available: \textcolor{black}{\url{https://en.wikipedia.org/wiki/Nmap}}

\bibitem{metasploit}
Metasploit Framework, Wikipedia. [Online]. Available: \textcolor{black}{\url{https://en.wikipedia.org/wiki/Metasploit}}

\bibitem{owaspzap}
OWASP ZAP, OWASP Foundation. [Online]. Available: \textcolor{black}{\url{https://www.zaproxy.org/}}

\bibitem{burpsuite}
Burp Suite, Wikipedia. [Online]. Available: \textcolor{black}{\url{https://en.wikipedia.org/wiki/Burp_Suite}}

\bibitem{sqlmap}
SQLmap, Wikipedia. [Online]. Available: \textcolor{black}{\url{https://en.wikipedia.org/wiki/Sqlmap}}

\bibitem{nessus}
Tenable, ``Nessus Vulnerability Scanner.'' [Online]. Available: \textcolor{black}{\url{https://www.tenable.com/products/nessus}}

\bibitem{mstmt}
Microsoft, ``Threat Modeling Tool,'' Microsoft Learn. [Online]. Available: \textcolor{black}{\url{https://learn.microsoft.com/en-us/azure/security/develop/threat-modeling-tool}}

\bibitem{threatdragon}
OWASP Threat Dragon, OWASP Foundation. [Online]. Available: \textcolor{black}{\url{https://owasp.org/www-project-threat-dragon/}}

\bibitem{mitreattack}
MITRE Corporation, ``MITRE ATT\&CK Framework.'' [Online]. Available: \textcolor{black}{\url{https://attack.mitre.org/}}

\bibitem{capec}
MITRE, ``CAPEC -- Common Attack Pattern Enumeration and Classification.'' [Online]. Available: \textcolor{black}{\url{https://capec.mitre.org/about/documents.html}}

\bibitem{misp}
MISP Project, ``Malware Information Sharing Platform \& Threat Sharing.'' [Online]. Available: \textcolor{black}{\url{https://www.misp-project.org/}}

\end{thebibliography}

